---
title: "MIR4435-2HG ASO design for exon 1"
author: "Brooke Morrison"
date: "2024-05-28"
output: html_document
---

In this document I am taking the Sfold output generated for exon 1 of MIR4435-2HG, filtering it based on the ASO pipeline from Zohaib, and saving it as a fasta file. 

```{r, message = FALSE}
# Loading packages
library(stringr)
library(Biostrings)
```

```{r}
# Loading in data
Sfold_unfiltered3 <- read.table("/Volumes/archive/diermeierlab/Brooke_MSc/data/MIR4435-2HG_general_output_exon1/output/oligo.txt", header = FALSE, sep = "", fill = TRUE)

# Labeling the rows
rownames(Sfold_unfiltered3) <- Sfold_unfiltered3[ , 1]
Sfold_unfiltered3 <- Sfold_unfiltered3[, -1]
head(Sfold_unfiltered3)
nrow(Sfold_unfiltered3) #431 of then (unfiltered)
```

```{r}
# Removing GGGG indicator
processed_sfold_3 <- Sfold_unfiltered3[ , -6]

# Going to add column labels
colnames(processed_sfold_3) <- c("Target position (start-end)", "Target sequence (5'-3')", "Antisense oligo (5'-3')", "GC content", "Oligo binding energy (kcal/mol")
head(processed_sfold_3)
```

```{r}
# Filtering for GC content 40-60%
processed_sfold_3 <- processed_sfold_3[processed_sfold_3[, 4] >= 40 & processed_sfold_3[, 4] <= 60, ]
nrow(processed_sfold_3) #203
```

```{r}
# Filtering for binding energy less than -8
processed_sfold_3 <- processed_sfold_3[processed_sfold_3[, 5] < -8, ]
head(processed_sfold_3)
nrow(processed_sfold_3) #35
```
There are 35 ASO options here with binding energies less than -8

```{r}
# Looking for diminishing and enhancing motifs
enhancing_motifs <- c("CCAC", "TCCC", "ACTC", "GCCA", "CTCT")
diminishing_motifs <- c("GGGG", "ACTG", "AAA", "TAA")

# Are any enhancing motifs present?
processed_sfold_3$EnhancingMotif_Present <- sapply(processed_sfold_3$`Antisense oligo (5'-3')`, function(seq) {
  any(str_detect(seq, enhancing_motifs))
})

# Are any diminishing motifs present?
processed_sfold_3$DiminishingMotif_Present <- sapply(processed_sfold_3$`Antisense oligo (5'-3')`, function(seq) {
  any(str_detect(seq, diminishing_motifs))
})

# Removing those with diminishing motifs present
processed_sfold_3 <- processed_sfold_3[processed_sfold_3$DiminishingMotif_Present == FALSE, ]
nrow(processed_sfold_3) #19
```
There are 19 ASOs after those with diminsihing motifs were removed

Are any of these also present in oligo_de.out file? - Going to load in this file, and use intersect function to see if any are present in both.

Column 1: starting target position
Column 2: ending target position
Column 3: target sequence (5p --> 3p)
Column 4: antisense oligo (5p --> 3p)
Column 5: GC content
Column 6: average unpaired probability for target site nucleotides
Column 7: binding site disruption energy (kcal/mol)

FILTER CRITERIA: ("<=": less than or equal to)
                 (">=": greater than or equal to)

 A) 40% <= GC % <= 60%;
 B) No GGGG in the target sequence;
 C) Average unpaired probability for target site nucleotides >= 0.5;
 D) For each peak in the accessibility profile that is above the threshold
    probability of 0.5, all sites targeted to this same peak are
    ranked by their average unpaired probability (the higher the better) and
    at most n sites are selected for each peak, where n is determined by
    max([width of peak/site length], 2);
 E) Among sites satisfying criteria A-D, the top 20 unique ones with
    the highest average unpaired probability are listed.

NOTE:
 i) The average unpaired probability is used in filter criteria C, D and E to
    cut down the number of reported sites in order to make the disruption
    energy calculation manageable on our web servers.
    
```{r}
oligo_de.txt.3 <- read.table("/Volumes/archive/diermeierlab/Brooke_MSc/data/MIR4435-2HG_general_output_exon1/output/oligo_de.txt", header = FALSE, sep = "", fill = TRUE)

colnames(oligo_de.txt.3) <- c("Starting target position", "Ending target position", "target sequence (5'-3')", "Anti-sense oligo (5'-3')", "GC content", "Average unpaired propbability for target site nucleotides", "Binding site disruption energy")

shared_ASOs.3 <- intersect(processed_sfold_3$`Antisense oligo (5'-3')`, oligo_de.txt.3$`Anti-sense oligo (5'-3')`)
shared_ASOs.3 # four of them

shared_ASO_df.3 <- processed_sfold_3[processed_sfold_3$`Antisense oligo (5'-3')` %in% shared_ASOs.3, ]

shared_ASO_df.3
```
There are four ASOs in the processed oligo.out (processed_sfold_3) data that are also in the oligo.de.txt file, have made a new object ("shared_ASO_df.3") that contains only the shared ASOs identified by both output files. They are targetting relatively different areas in the exon, and all have great binding energies.

```{r}
# Save the output as a fasta file (using Biostrings)

# Putting it into a seq ID, and sequence format
ASO_df.3 <- data.frame(
  names = c("Exon1_1", "Exon1_2", "Exon1_3", "Exon1_4"),
  sequences = c("GGAGAACTTTGTTGTGATGG", "ATGAGGGAGAACTTTGTTGT", "CATGAGGGAGAACTTTGTTG", "TGTATATGATTCTTCACGGC")
)

# Converting df to DNAStringSet object
sequences <- DNAStringSet(ASO_df.3$sequences)
names(sequences) <- ASO_df.3$names

# Write DNAStringSet object to FASTA file
writeXStringSet(sequences, filepath = "MIR44352HG_exon1_ASOs.fasta")
```

